---
title: "Homework 1"
author: "Aleksei Parm"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    code_folding: show
---

```{r, include = FALSE}
#setwd("~/University/statistical-learning/hws/hw1") #set working directory
```



# About

The homework is about exploring [`Carseats_mod.csv`](https://github.com/aparm/statistical-learning/blob/main/hws/hw1/Carseats_mod.csv) data set. 

[GitHub](https://github.com/aparm/statistical-learning/tree/main/hws/hw1)



# Data

## Info about variables:

- Sales - Unit sales (in thousands) at each location
- CompPrice - Price charged by competitor at each location
- Income - Community income level (in thousands of dollars)
- Advertising - Local advertising budget for company at each location (in thousands of dollars)
- Population - Population size in region (in thousands)
- Price - Price company charges for car seats at each site
- ShelveLoc - A factor with levels Bad, Good and Medium indicating the quality of the shelving location for the car seats at each site
- Age -Average age of the local population
- Education -Education level at each location
- Urban - A factor with levels No and Yes to indicate whether the store is in an urban or rural location
- US - A factor to indicate whether the store is in the US or not



# Libraries

```{r setup}
#install.packages("tidyverse")
#install.packages("GGally")

library(tidyverse)
library(GGally)
```



# Exercises:

## Reading data

1. Check visually the file for presence of missing values. Use `read_delim` command (indicating the symbol for missing values if needed) to read in the data set `Carseats_mod.csv`. Make sure that all variables that, according to description, should be numeric are numeric after reading data in.

```{r}
carseats = read_delim("Carseats_mod.csv", delim=",", na = c(".", "NA"))
```

> Missing values in the file are presented as `.` symbol.  

<br>

2. Use `summary()` function to look at summary information about variables. Do you see anything which may indicate data errors? What? How many missing values are in the data set?

```{r}
summary(carseats)
```

> There are 3 missing values in the data set.



## Data cleaning

3. Use `na.omit()` command to define a new data set `carseats2` where rows with missing values in the original data are left out

```{r}
carseats2 <- na.omit(carseats)
summary(carseats2)
```

> New data set contains 397 rows

<br>

4. Use `select()` command to select columns `Sales` and from `Price` to `US` from `carseats2` and apply `ggpairs()` function to the the result. Are any possible data errors visible from the output?

```{r message=FALSE, warning=FALSE}
carseats2 %>%
  select(Sales, Price:US) %>%
  ggpairs()
```

> `US` factor has 3 different levels: `Yes`, `No`, `yes`.



## Data transformation

5. Use `mutate()` command with `if_else()` to correct data error in one of the factor variables, store the corrected data with the same name `carseats2`

```{r}
levels(as.factor(carseats2$US))

carseats2 <- carseats2 %>%
  mutate(US = if_else(US == "yes", "Yes", US))

levels(as.factor(carseats2$US))
```

> Now `US` factor has 2 levels: `Yes`, `No`.



## Summary statistics

6. Use `summarize()` command from `dplyr` package to compute a table with mean sales amount for each combination of `Urban` and `ShelveLoc`

```{r message=FALSE, warning=FALSE}
carseats2 %>%
  group_by(Urban, ShelveLoc) %>%
  summarize(MeanSales = mean(Sales))
```



## Data visualization

7. Use `ggplot()` with `geom_point` to produce a scatter plot of `Price` (on x-axis) and `Sales`, color points according to shelve location. Discuss the graph - is it showing expected relationships between variables?

```{r}
ggplot(carseats2, aes(Price, Sales, color = ShelveLoc)) +
  geom_point()
```

> The scatter plot is showing relationships:
>
> - between `Price` and `Sales` (Car seats with lower price usually have more sales)
> - between `ShelveLoc` and `Sales` (Car seats with better shelving location usually have more sales)


